An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
tweak
topk app

\begin{definition}
  \label{def:calibrated}
  Let original loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
\end{definition}
It is well-known that calibration implies \emph{consistency}, in the same sense as~\citet{agarwal2015consistent}.
Given feature space $\X$, fix a fixed distribution $D \in \Delta(\X\times\Y)$.
Let $L^*$ be the best possible expected $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
Then $(L,\psi)$ is consistent if a sequence of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to $L^*$, then the $\ell$-loss of $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to $\ell^*$.
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.

@@

We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
  For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let us first simplify $L'$:
  \[
    L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
  \]
  Next, we observe that $L'$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.
  To see this,
  \begin{align*}
    L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
    &= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
    &= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
  &= L'(u)_y
  \end{align*}

  We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  This follows from the observation that if $u_{[1]} > 1$, the the loss only decreases by reporting $u_{[1]} = u_{[n]} + 1 = 1$.
  For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
  \begin{align*}
  	\inprod{p}{L'(u)} &= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}
  To verify this, observe the first term of the summation comes from the probability that you incorrectly did not have $u_y$ in the top-$k$, and the second term comes from the $(k+1)^{th}$ term being added when $u_y$ is correctly in the top $k$ (and therefore not an averaged term, since $u_y \in [0,1]$.)
  The $1$ is pulled out from the average, and $\inprod{p}{u}$ comes from the $-u_y$ term form each $y$ by linearity of expectation.

  Since, for $u_{[j]}$ and $j \geq k+2$, the expected loss only decreases by increasing $u_{[j]}$ as much as possible, we increase them to $u_{[k+1]}$.
  By our invariance in the $\ones$ direction, we can re-shift our values if necessary, setting $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.

  Substituting these values into the expected loss, we now observe a linear objective over the hypercube, which is known to be optimal on some corner of the hypercube, so the rest of $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
  Therefore, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$.
  Moreover, for this optimal $u$, observe $\|u\|_{1} \leq k$ since we have $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$, so we observe that $u \in \U$ as well.
\end{proof}

>>>>

\begin{definition}
  \label{def:calibrated}
  Let original loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
\end{definition}
It is well-known that calibration implies \emph{consistency}, in the following sense (cf.~\cite{agarwal2015consistent}).
Given feature space $\X$, fix a fixed distribution $D \in \Delta(\X\times\Y)$.
Let $L^*$ be the best possible expected $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
Then $(L,\psi)$ is consistent if a sequence of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to $L^*$, then the $\ell$-loss of $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to $\ell^*$.
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.

@@

We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
  For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let us first simplify $L'$:
  \[
    L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
  \]
  Examining this expression, observe that
  if $u_i < u_{[k+1]}$, then $(u-e_y)_i < (u-e_y)_{[k]}$, as only one entry changes (and decreases) from one to the other.
  Hence, if $u_i < u_{[k+1]}$, then the index $i$ will not appear in the summation for any $y$, and thus only influences the loss when $y=i$.
  Now consider the loss of $u'$ which is the same as $u$ but with $u_i' = u_{[k+1]}$.
  Then $L'(u')_y \leq L'(u)_y$ when $y=i$ and $L'(u')_y = L'(u)_y$ when $y\neq i$.
  From this we see that without loss of generality, $u_{[k+1]} = \cdots = u_{[n]}$.

  Next, we show that $L'$ is invariant in the $\ones$ direction:
  \begin{align*}
    L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
    &= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
    &= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
  &= L'(u)_y~.
  \end{align*}
  Thus, without loss of generality, we now have $u_{[k+1]} = \cdots = u_{[n]} = 0$.

  Let $u_{[1..j]} := \sum_{i=1}^j u_{[i]}$.
  We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
  \raf{Only hole left is the following line.  I think it might not be true.}
  This follows from the observation that if $u_i > 1$, the the loss only decreases by reporting $u_i = 1$.
  For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
  \begin{align*}
  	\inprod{p}{L'(u)} &= 1 - \inprod{p}{u} + \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} \right) = 1 + \inprod{\tfrac 1 k \ones - (1+\tfrac 1 k)p}{u}~.
  \end{align*}
  As a linear objective over the hypercube, $\inprod{p}{L'}$ must be optimal on some corner of the hypercube, so we conclude $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
  In particular, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$, and since
  Moreover, there is an optimal $u$ with $\|u\|_{1} \leq k$ since we have $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$, so some $u \in \U$ must be optimal as well.
\end{proof}
